---
title: "task4_analysis"
output: html_document
date: "2023-03-13"
---

```{r}
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(dplyr)
library(stringr)
```

# Need to do a full join of the data:

```{r}
test_join_2015_2016 <- full_join(clean_data_2015, clean_data_2016)

test_join_2015_2016 # 6,889 x 102

full_candy <- full_join(test_join_2015_2016, clean_data_2017)

full_candy # 9,349 x 106
```

# Assumptions made and things to note:

# Mary Janes 2 is the brown globs line. When checking for most popular candy, we need to add these together. 





# 1 - What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

Need to count all instances of Joy, Meh and Despair

```{r}
full_candy %>% count(butterfinger, name = "Ratings", sort = TRUE)

# JOY = 6093
# DESPAIR = 1403
# NA = 1088
# MEH = 765
# OVERALL = 4690
```

```{r}
sum_joy <- sum(str_count(full_candy, "JOY")) # 277,361
sum_meh <- sum(str_count(full_candy, "MEH")) # 79,418
sum_despair <- sum(str_count(full_candy, "DESPAIR")) # 274,993

sum_joy + sum_meh + sum_despair # Total is 631,772
```

# 2 - What was the average age of people who are going out trick or treating?

```{r}
going_out_yes <- subset(full_candy, going_out == "Yes" & age >= 1 & age <= 100, select = c(age, going_out))
going_out_yes  

going_out_no <- subset(full_candy, going_out == "No" & age >= 1 & age <= 100, select = c(age, going_out))
going_out_no

going_out_yes_clean <- na.omit(going_out_yes) # This reduces the rows from 981 to 925.
going_out_yes_clean # I considered removing data where the age was 4 etc, but these could have been completed by a parent so make sense to keep

going_out_no_clean <- na.omit(going_out_no) # This reduces the rows from 8,258 to 7,880.
going_out_no_clean

mean(going_out_yes_clean$age, na.rm = TRUE) # 252 96 years (oh no, lesson learned here to check data first)
mean(going_out_no_clean$age, na.rm = TRUE) # 39.93 years 

sort(going_out_yes_clean$age) # There are a few crazy inputs here, 200,587, 1,000 etc. We will choose a range of 1 year to 100 years. 

mean(going_out_yes_clean$age, na.rm = TRUE) # 35.02 years (yay)
mean(going_out_no_clean$age, na.rm = TRUE) # 39.93 years 
```

# 3 - What was the average age of people who are not going trick or treating?

# I answered this above. 

# 4 - For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
# full_candy %>% count(butterfinger, name = "Ratings", sort = TRUE)



sort(apply(candies_only, 2, table))

# Need to have all of the above info in one table to be able to sort the data

# Group by count using dplyr
#agg_tbl <- full_candy %>% group_by(department) %>% 
#  summarise(total_count=n(),
#            .groups = "JOY")
#agg_tbl

# Convert tibble to df
#df2 <- agg_tbl %>% as.data.frame()
#df2

#candy_names <- 

#candy_order[order(candy_only$JOY),]

# Need to pivot the data so that we have JOY, MEH and DESPAIR as the column headings with the counts as the data

candies_only <- select(full_candy, -internal_id, -timestamp, -age, -going_out, -year, -hugs_actual_physical_hugs, -country, -gender)

candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:97),
                                   names_to = "candy",
                                   values_to = "feelings")

# Select columns
dt <- candies_only_long[, c("feelings", "candy")] %>% 
  filter(feelings == "JOY")

# Unlist the data frame
dt_vec <- unlist(dt)

sort(dt_vec)

# Count the number
table(dt_vec)

candies_grouped <- candies_only_long %>% 
  group_by(candies)  

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "JOY") %>% 
  arrange(desc(n))

# Count the number
table(sort(dt_vec))



full_candy %>%
  group_by("JOY", "MEH", "DESPAIR") %>%
  summarize(candies_only_long = n())

ratings <- c("JOY", "MEH", "DESPAIR")


candies_grouped <- candies_only_long %>% 
  group_by(candy)  
candies_grouped <- candies_only_long %>% 
  count(candy, feelings) 

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "JOY") %>% 
  arrange(desc(n))






# Full code to work below:


candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:97),
                                   names_to = "candy",
                                   values_to = "feelings")


candies_grouped <- candies_only_long %>% 
  group_by(candy)  
candies_grouped <- candies_only_long %>% 
  count(candy, feelings) 

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "JOY") %>% 
  arrange(desc(n))

# 7,589 for any full sized candy bar, if this isn't allowed then 7,369 for reese's peanut butter cups

candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:97),
                                   names_to = "candy",
                                   values_to = "feelings")


candies_grouped <- candies_only_long %>% 
  group_by(candy)  
candies_grouped <- candies_only_long %>% 
  count(candy, feelings) 

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "MEH") %>% 
  arrange(desc(n))

# 1,570 lollipops 

candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:97),
                                   names_to = "candy",
                                   values_to = "feelings")


candies_grouped <- candies_only_long %>% 
  group_by(candy)  
candies_grouped <- candies_only_long %>% 
  count(candy, feelings) 

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "DESPAIR") %>% 
  arrange(desc(n))

# 7,905 broken glow stick 
```



# 5 - How many people rated Starburst as despair?

```{r}
candies_grouped %>% 
  select(candy, feelings, n) %>% 
  filter(candy == "starburst", feelings == "DESPAIR")
```




# For the next three questions, count despair as -1, joy as +1, and meh as 0.

# 6 - What was the most popular candy bar by this rating system for each gender in the dataset ?
What was the most popular candy bar in each year?
What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
# We removed gender earlier but need to add it back in for question 6:

full_candy %>% count(gender = "Male", sort = TRUE)

full_candy %>% count(gender = "Female", sort = TRUE)











candies_only <- select(full_candy, -internal_id, -timestamp, -age, -going_out, -year, -hugs_actual_physical_hugs, -country)

candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:98),
                                   names_to = "candy",
                                   values_to = "feelings")



candies_grouped <- candies_only_long %>% 
  group_by(candy)  
candies_grouped <- candies_only_long %>% 
  count(candy, feelings) 

candies_grouped_male %>% 
  select(candy, feelings, n) %>% 
  slice_max(gender == "Male" & feelings == "DESPAIR") %>% 
  arrange(desc(n))



candies_grouped <- candies_only_long %>% 
  group_by(candy)  
candies_grouped <- candies_only_long %>% 
  count(candy, feelings) 

candies_grouped_female %>% 
  select(candy, feelings, n) %>% 
  slice_max(gender == "Female" & feelings == "DESPAIR") %>% 
  arrange(desc(n))




```

```{r}
# changing the rating column to 1, -1 and 0

#rating_count_new <- candies_grouped %>% 
#  mutate(rating = case_when(rating == "DESPAIR" ~ "-1",
#                            rating == "JOY" ~ "1",
#                            rating == "MEH" ~ "0",
                            TRUE ~ rating))

# converting rating column to numeric

#rating_count_new_numeric <- rating_count_new %>% 
#  mutate(rating = as.numeric(rating))

# calculating the sum of rating for each candy

#rating_count_new_numeric_sum <- rating_count_new_numeric %>% 
#  filter( rating == -1 | rating == 1 | rating == 0) %>% 
#  mutate( new_rating = rating * n ) %>% 
#group_by(candies) %>% 
#summarise(sum_rating = sum(new_rating))

# reordering the table descendant to see the highest rated candy

#rating_count_new_numeric_sum_ordered <- rating_count_new_numeric_sum %>% 
#  arrange(desc(sum_rating))

#rating_count
#rating_count_new
#rating_count_new_numeric
#rating_count_new_numeric_sum
#rating_count_new_numeric_sum_ordered









# selecting for candies, gender, year and country

candies_only_q6 <- select(full_candy, -age, -going_out)

# changing order so gender column is first

candies_only_q6 <- candies_only_q6 %>% 
  relocate(gender, .before = butterfinger) %>%
   relocate(year, .after = gender) %>% 
   relocate(country, .after = year)
  
# new df with gender, feelings and candy

candies_only_q6_long <- candies_only_q6 %>% 
  pivot_longer(
              cols = c(4:100),
              names_to = "candy",
              values_to = "feelings")

# duplicated feelings column

candies_only_q6_long$rating = candies_only_q6_long$feelings

# replaces feelings with rating 1, -1 and 0

rating_count_new <- candies_only_q6_long%>% 
  mutate(rating = case_when(rating == "DESPAIR" ~ "-1",
                            rating == "JOY" ~ "1",
                            rating == "MEH" ~ "0",
                            TRUE ~ rating))

# changes rating to numeric
rating_count_new_numeric <- rating_count_new %>% 
  mutate(rating = as.numeric(rating))





```

# Rating Score = Number of JOY - Number of DESPAIR

```{r}
# Using the count for Starburst
# candies_grouped %>% 
#  select(candy, feelings, n) %>% 
#  filter(candy == "starburst", feelings == "DESPAIR")



```

# Third attempt as getting a lot of error messages above. Felt easier to start again:

```{r}
## reassigning and re-positioning year, country and gender to the data set

full_candy_new <- select(full_candy, - timestamp, -age, -going_out, 
                       -hugs_actual_physical_hugs, -internal_id) %>% 
  mutate(year = as.character(year)) %>% 
  relocate(year, .after = take_5) %>% 
  relocate(gender, .before = year) %>% 
  relocate(country, .after = year)

full_candy_new

# pivoting the full_candy_new table 

full_candy_new_long  <- pivot_longer(full_candy_new,
                                   cols = c(1:98),
                                   names_to = "candies",
                                   values_to = "rating")

# counting the rating and converting gender to NA where the answers were not wrong

rating_count_full_candy <- full_candy_new_long %>% 
  count(candies, rating, year, gender,country)  %>% 
  mutate (gender = case_when(gender == "I'd rather not say" ~ NA_character_ ,
                              gender == "Other" ~ NA_character_,
                              gender == "NA" ~ NA_character_,
                              TRUE ~ gender))

# converting the rating column to -1. 1, and 0 

rating_count_full_candy_new <- rating_count_full_candy %>% 
  mutate(rating = case_when(rating == "DESPAIR" ~ "-1",
                            rating == "JOY" ~ "1",
                            rating == "MEH" ~ "0",
                            TRUE ~ rating))

# converting rating column to numeric

rating_count_full_candy_new_numeric <- rating_count_full_candy_new %>% 
  mutate(rating = as.numeric(rating))

# calculating the sum of rating for each candy per gender

popular_candy_by_gender <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies, gender) %>%
summarise(sum_rating = sum(new_rating)) 
  

# reordering the table descendant to see the highest rated candy

male_popular_candy <- popular_candy_by_gender %>% 
 filter(gender == "Male") %>% 
arrange(desc(sum_rating))

female_popular_candy <- popular_candy_by_gender %>% 
 filter(gender == "Female") %>% 
arrange(desc(sum_rating))

male_popular_candy # 1,584 for any full sized candy bar

female_popular_candy # 875 for any full sized candy bar
```

# 7 - What was the most popular candy bar in each year?

```{r}
popular_candy_by_year <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies, year) %>%
summarise(sum_rating = sum(new_rating)) 

popular_candy_by_year

popular_candy_2015 <- popular_candy_by_year %>% 
 filter(year == "2015") %>% 
arrange(desc(sum_rating))

popular_candy_2016 <- popular_candy_by_year %>% 
 filter(year == "2016") %>% 
arrange(desc(sum_rating))

popular_candy_2017 <- popular_candy_by_year %>% 
 filter(year == "2017") %>% 
arrange(desc(sum_rating))

popular_candy_2015 # 4,603 for any full sized candy bar, 4,375 for reeses peanut candy bar

popular_candy_2016 # 1,037 for any full sized candy bar, 920 for kitkat

popular_candy_2017 # 1,542 for any full sized candy bar, 1,403 for reeses peanut candy bar

# I want to check the answer is correct:
# Overall rating for any full sized candy bar is:

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  filter(candy == "any_full_sized_candy_bar", feelings == "JOY") # 7,589

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  filter(candy == "any_full_sized_candy_bar", feelings == "DESPAIR")# 407 so JOY - DESPAIR = 7,182

# Total of popular candy = 4,603 + 1,037 + 1,542 = 7,182 - YAY!
```
# 8 - What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
# For USA:

popular_candy_by_country <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 | rating == 0 ) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies, country) %>%
summarise(sum_rating = sum(new_rating)) 

popular_candy_by_country

popular_candy_USA <- popular_candy_by_country %>% 
 filter(country == "USA") %>% 
arrange(desc(sum_rating))

popular_candy_USA # 2,172  for any full sized candy bar, 1,983 for reeses peanut butter cups

popular_candy_UK <- popular_candy_by_country %>% 
 filter(country == "UK") %>% 
arrange(desc(sum_rating))

popular_candy_UK # 33 for any full sized candy bar, lindt truffle, rolos, toblerone

popular_candy_Canada <- popular_candy_by_country %>% 
 filter(country == "Canada") %>% 
arrange(desc(sum_rating))

popular_candy_Canada # 252 for any full sized candy bar, 230 for kitkat


```







